---
title: "Spotify top 2010-2019 analysis"
output: html_notebook
---
# Loading libraries
```{r}
library(MASS)
library(ggplot2)
library(gridExtra)
library(digest)
library(hrbrthemes)
library(GGally)
library(viridis)
library(fmsb)
```

# Initializing data
```{r}
top10s <- fread("../top10s.csv")
names(top10s)[names(top10s) == "top genre"] <- "genre"
names(top10s)[names(top10s) == "V1"] <- "ID"
top10sATTR <- subset(top10s, select = -c(ID, title, genre, artist, pop))
```

# Popularity scatter plot
```{r}
orderedPOP <- top10s[order(pop),]
orderedPOP$ID <- seq.int(nrow(s))
ggplot(orderedPOP, aes(x=ID, y=pop)) + geom_point(size=2, shape=23)
```

# Histogram of genres
```{r}
ggplot(top10s, aes(x=genre, color=genre, fill=genre)) + stat_count() + theme(axis.text.x=element_blank())
```
The *dance pop* genre is dominant followed by *pop* and *canadian pop*. The other genres are minor.

# Boxplots of popularity attribute for each genre
```{r}
ggplot(top10s) + geom_boxplot(aes(x=top10s$genre, color=genre, fill=genre, y=top10s$pop)) + labs(x="genre", y="popularity") + theme(axis.text.x=element_blank())
```
From here we can see that the most successful genres are *brostep* and *electro*. As these genres are minor compared to *dance pop*, it is difficult to say if there is some kind of rule behind their popularity.

# Correlation between attributes
```{r}
ggpairs(top10s, columns = 6:15)
```
Here we will only discuss attributes with at least 0.1 correlation value.

So top 4 most popularity-influencing attributes are:

1. *dB* (loudness) with 0.157
2. *dnce* (danceability) with 0.116
3. *dur* (duration) with -0.104

We also can see that:

* *dB* is correlated with *val* (valence), *nrgy* (energy) and already mentioned *dnce*
* *dnce* as well is correlated with *val* and *nrgy*, but also with *bpm* and *dur*
* *dur* is correlated with *nrgy*, *dnce* and *dB*

So after this rough analysis we could say that the most important attributes for popularity are:

*dB*, *dnce*, *dur*, *val*, *nrgy* and *bpm*
